A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-11-07 imaging_demo/ms.tex
Move software architecture section to appendix
From cb584b3 (review):
=====================

In ms.tex,

(*) Move section 3.2 (software architecture) to appendix
(*) Remove section 3.1 heading "Data flow"
(*) Add a sentence at end of section 3 to refer the interested reader
to the appendix
(*) Update abstract
(*) Few minor changes

% Abstract of the paper
\begin{abstract}
Modern radio telescopes are favoring densely packed array layouts consisting of 
large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of 
traditional correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep 
cost for realizing the full imaging potential of these powerful instruments. 
Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), 
we present the first software demonstration of a generalized direct imaging 
algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
advantage of the multiplication-convolution theorem of Fourier transforms. Not 
only does it bring down the cost for dense layouts to 
$\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregularly arranged 
heterogeneous antenna arrays. EPIC is highly modular and parallelizable, 
implemented in object oriented Python, and publicly available. We have verified 
the images produced to be equivalent to those produced using traditional 
techniques to within a precision determined by coarseness of gridding. We have 
also validated our implementation on data observed with the Long Wavelength Array 
(LWA). Antenna layouts with a dense filling factor consisting of a large number 
of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of 
Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain 
significant computational advantage by deploying EPIC. Inherent availability of 
calibrated time-domain images on timescales roughly equal to the writeout 
timescale of the digitizer and vastly lower I/O bandwidth relative to visibility 
based systems will make it a prime candidate for transient searches of Fast Radio 
Bursts (FRB) as well as planetary and exoplanetary phenomena. 
\end{abstract}

@@

For certain classes of radio arrays there is an alternative to the FX correlator
that can lower the computational burden by directly performing a spatial fast
Fourier transform (FFT) on the electric fields measured by each antenna in the
array at each time step, removing the cross-correlation step. This relieves the
computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
$\sim\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of grid points in the 
Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often 
referred to as a ``direct imaging'' correlator because it eliminates the 
intermediate cross-correlation data products of the FX and XF correlators, but 
instead directly forms images from the electric field measurements.

@@

\subsection{Data Flow}

@@

\subsection{Software Architecture}\label{sec:software-modules}

\subsubsection{Antenna Module}

\subsubsection{Interferometer Module}

\subsubsection{Antenna Array Module}

\subsubsection{Interferometer Array Module}

\subsubsection{Image Module}

@@

The number of complex multiplications and additions in the spatial Fourier 
transform implemented via Fast Fourier Transform \citep[FFT;][]{coo65} is 
$\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of 
pixels on the grid, the factor 4 accounts for increase in number of pixels as a 
result of zero-padding before spatial Fourier transform, and $\beta$ is a 
constant that depends on the implementation of twiddle FFT algorithms 
\citep{bri74}. In our study, we set $\beta=5$, a 
value\footnote{http://www.fftw.org/speed/method.html} much more conservative 
than was indicated in \citet{mor11}. We set the number of complex 
multiplications in the X-engine in visibility based imaging to 
$\Nant(\Nant-1)/2$.

@@

% \appendix

>>>>

% Abstract of the paper
\begin{abstract}
Modern radio telescopes are favoring densely packed array layouts with 
large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of 
traditional correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep 
cost for realizing the full imaging potential of these powerful instruments. 
Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), 
we present the first software demonstration of a generalized direct imaging 
algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
advantage of the multiplication-convolution theorem of Fourier transforms. Not 
only does it bring down the cost for dense layouts to 
$\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregularly arranged 
heterogeneous antenna. EPIC is highly modular and parallelizable, implemented in 
object oriented Python, and publicly available. We have verified the images 
produced to be equivalent to those produced using traditional techniques to 
within a precision determined by coarseness of gridding. We have also validated 
our implementation on data observed with the Long Wavelength Array (LWA). Antenna 
layouts with a dense filling factor consisting of a large number of antennas such 
as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and 
Canadian Hydrogen Intensity Mapping Experiment will gain significant 
computational advantage by deploying EPIC. Inherent availability of calibrated 
time-domain images on digitizer writeout timescales and vastly lower I/O 
bandwidth relative to visibility based systems will make it a prime candidate 
for transient searches of Fast Radio Bursts (FRB) as well as planetary and 
exoplanetary phenomena. 
\end{abstract}

@@

For certain classes of radio arrays there is an alternative to the FX correlator
that can lower the computational burden by directly performing a spatial Fast
Fourier Transform \citep[FFT;][]{coo65} on the electric fields measured by each 
antenna in the array at each time step, removing the cross-correlation step. 
This relieves the computational scaling from the harsh $\Nant^2$ to the more 
gentle envelope of $\sim\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of 
grid points in the Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This 
architecture is often referred to as a ``direct imaging'' correlator because it 
eliminates the intermediate cross-correlation data products of the FX and XF 
correlators, but instead directly forms images from the electric field 
measurements.

@@

% \subsection{Data Flow}

@@

A high level software architecture of the EPIC package is described in the 
appendix \S\ref{sec:software-modules} for the interested reader.

@@

The number of complex multiplications and additions in the spatial Fourier 
transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ 
where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for 
increase in number of pixels as a result of zero-padding before spatial Fourier 
transform, and $\beta$ is a constant that depends on the implementation of 
twiddle FFT algorithms \citep{bri74}. In our study, we set $\beta=5$, a 
value\footnote{http://www.fftw.org/speed/method.html} much more conservative 
than was indicated in \citet{mor11}. We set the number of complex 
multiplications in the X-engine in visibility based imaging to 
$\Nant(\Nant-1)/2$.

@@

\appendix

\section{Software Architecture}\label{sec:software-modules}

\subsection{Antenna Module}

\subsection{Interferometer Module}

\subsection{Antenna Array Module}

\subsection{Interferometer Array Module}

\subsection{Image Module}
